\chapter{All-in is not a value bet}

% Last updated: 20250416
% CHAPTERLENGTH: 2

% FIXME: put this chapter into the Bet-sizing stuff and not
% a separate chapter?

One bad play I've seen from a few people is pushing all-in
with strong but non-nut hands. The bet will usually win the
pot, because worse hands (that you want a call from, because they
lose) fold to such a strong bet. But if you do run
into a super strong hand, you've already gone all-in and given
away your chips.

\textbf{At the tables} I hold \Jh\tens\ and I've got quads on this
finalboard: \tenh\tenc\eigh\Qd\tend\ .

Very nice quads on the river for me. My opponent, first to act,
massively overbets all-in. He bet 90,000 into a 16,000 pot (he had
140,000, I had 90,000) and I make the easiest of calls. He shows Q9.

There's four hands that can beat his very strong TTTQQ full house.
From stronger to strongest they are: KK (TTTKK), AA (TTTAA), QQ
(better full house QQQTT) and Tx (TTTT quads). I'll call with Qx for a
probable split, but any other hands besides the winners above I'll
fold.

So this bet acheives the exact opposite of what a river bet wants.
It can't fold better hands, and it can't get a call from weaker hands.
Q9 is a strong hand on a board of TT8QT, but it's not the best hand,
and if you bet all-in with it you'll only get called by a better hand.

You can see the same thing yourself if you shove all-in with a
one-card King-high flush every time there's four flush cards on the
board. You'll get the Jack-high flushes to fold, which you
wanted a losing call from, and if the Ace-high flush is there,
you've given away your chips.

Of course you sometimes bet all-in without the nuts, depending on
your chip stack and the size of the pot and what hand you expect
your opponent to have and the reason for your bet. But overbetting
all in with a strong hand is usually wrong.

It's probably hard for Q9 not to lose it all in this hand. If he bets
the pot, 16,000 on the river, I'll probably just reraise all in and hope
he's got a Queen. But if I read him as weak I might just reraise to
40,000 and then he gets to call and he saves a lot of chips.
