\chapter{Australian Poker Weekly}

% Last updated: 20200418
% Only a short intro paragraph so this chapter stays at 2 pages

Australian Poker Weekly was a newspaper devoted to poker in Australia
that came out every Friday. It ran from January 2009 till June 2011.
I wrote over 120 articles in that period. These were so good that
I've collected and published them as a second book.
Here's the article from Issue 4 of February 2009.

\section{Considering a Big Laydown}

It's early January and I'm playing in a small weeknight \$10 NPL game,
close to where I live. The game is small (typically 20 to 35 runners) with
\$500 in guaranteed prizemoney, with the top 3 finishers getting \$250, \$150
and \$100. I'm defending champion, having won the same tournament
the week before.

It's the sixth hand of the night, blinds are still 25/50 but
my 2,500 start stack is now at 1,650. I'm the shortest stack at the
table with most players at the starting stack. I'm in the big blind and
in a hand with 2 limpers and the small blind completing, I look down to
find two red Aces and I raise to 500 straight. The two limpers get out
of the way, the small blind calls.

I always like playing AA in a raised pot against one opponent only.
Often this ensures he has a hand that plays badly against AA, such as
an underpair or AK or AT suited. Without a preflop raise, good
cracking hands as weak as 54 suited can see a flop, and spike a huge
hand on a flop like J55.

In this hand, I've built the pot to 1,100 and I'm up against one
opponent, the small blind, who's out of position for the remaining
betting rounds. I couldn't be in better shape, unless everyone had
folded and I got a 150 profit straight away.

The flop comes down JT3 rainbow. The small blind does a funny motion
where he gets about 650 in chips in his hand, advances it into the
table, withdraws his hand and his chips and finally checks. This is
no regulation insta-check, and I read this as a ``weak means strong''
tell. I think the small blind has hit this flop hard.

Flops containing both a Jack and a Ten are likely to be good flops
for calling hands, except missed underpairs like 22 to 99. I bet
500 after the small blind's strange action, and he check raises me
the minimum to 1,000.

Whoa. I have a really strong feeling that I'm behind here, but can
I get away from AA? The only 2 pair holding he could have that makes
sense is JT--- he's not calling 500 preflop with T3 or J3. Set hands
are JJ, TT and 33. He could have a worse overpair like QQ or KK,
but I'd expect a reraise preflop with those holdings.

Possibly he has AJ or KJ for top pair, good kicker, but I'd
expect a leading bet with such a hand, not a check-raise. Maybe he
has 89 or KQ for an open ended straight draw, but I think he'd just
call my 500 flop bet, and not check-raise. If he's check-raising
as a semibluff, why not check-raise me all in?

I now have 3 indications of small-blind strength: his preflop call
of a very decent raise, his strange chip motions on the flop and his
check-raise the minimum amount. The guy looks a straightforward player
who wouldn't go overboard just with top pair top kicker. If I fold
to this check-raise, I'll still have 650, very short
against all the stacks but at least keeping me in the tournament.

The small blind himself has had 2 indications of strength from
me--- a big preflop raise and a solid flop continuation bet of just
under half the pot and just under half my stack. Really he's not hoping
or expecting me to fold with his check raise.

Reasoning that ``it's a little \$10 NPL tournament'' I reraise my
last 150 in chips and the small blind calls and shows JJ for top set.
I'm now a 92\% chance to be out of this tournament and that's just where
I am two cards later.

You can play freerolls and \$10 games all week and not see a dent
in your wallet, but if you're not playing good poker in them you're
just picking up cards and shifting chips. Many of the players will
be taking a \$10 game lightly--- most of your advantage in the game
is treating it seriously and playing the best poker that you can.
Look for tells, put people on hands, size your bets well and
play proper poker. Treat any game you play with respect.

If it helps, imagine that your superior poker skills are giving
you that \$150 second prize with a shot at first. A stupid poker
mistake on your part won't cost you your \$10 buyin, it'll cost
you that \$150 second prize that should be yours against this
field of beginners.

A key skill in poker is trusting your reads and acting on them.
It's great to spot tells, but if you don't have the guts to follow
through on them you're wasting effort even looking for them. It's
great to spot weakness tells and make a hero call against a big bluff.
It's much harder to make hero folds after spotting strength tells,
especially when you have a strong hand yourself.
In this hand, my reading skills were great, but I lacked the guts to act
on my read and throw away my powerful AA.

In No Limit, where a single hand can eliminate you or cripple you, the key
question is ``Am I beaten?''. If the answer is ``Yes'', and you don't have
the outs to continue in the hand, the correct play is to fold.

Imagine the respect I'd have got from the table if I'd looked at the
small blind after that check-raise, said ``both your cards are
Jacks, I fold'' and thrown Aces into the muck face up!. I think the
small blind would have turned over his Jacks to show I'd made a great
laydown. The feeling of making a great laydown like that,
and the amazement from the other players would have been awesome, almost as
good as pocketing that \$150 second prize that should be mine.

