\chapter{Breaks}

% Last updated: 20200428

Every few rounds of a tournament there will be a 15 minute break.
You'll have to leave the poker table and playing area
and do other things for a while. The venue would like to sell
a few drinks at this point in proceedings, if you feel like one, grab one.
Don't play the poker machines during the break. Just don't.

If you're hungry or low on energy a chocolate bar might work
for you. I usually have a good break if I have a big stack
and a bad break if I'm short stacked and in tournament trouble.

\section*{Get some rest}

For a while I tried reading a Poker book during a break, but I
found that I was already tired from the tournament and I was much better
off resting my brain rather than trying to pick up new ideas.
Pubs and Clubs aren't great reading environments. Half an hour before a
tournament starts is a great time to read poker strategy books,
tournament breaks should be used as a complete break from the game.

\section*{Catch up with your poker friends}

If you see some poker pals in the break, ask them how they're
going and how many chips they have. Listen to a bad beat story
or two. If asked, tell them honestly how
many chips you have and how you think you're playing tonight.

Some of my poker pals are smokers so I'll often sneak outside and
chat to them. It can be good to get some fresh air (the outside
air, not the secondhand smoke) and temporarily reconnect with
the real world outside the poker tournament.

My poker pals are a great source of information on players
at the other tables, other games going on this week, home games
and everything else in the poker world. Many of my poker pals
have been tournament directors and some even helped to set up
poker games and indepedent rooms. I always
enjoy talking poker with the guys, and if you keep your ears open
you can learn a lot.

\section*{After the break}

An interesting phenomenon is that once play resumes after a break,
a lot of players will bustout very quickly. The tournament can
lose one-third or more of the remaining players in the first 10
hands after the break. The blinds have gone up to new highs and
all the players who were shortstacked before the break are
even more shortstacked now. There'll be quite a few players with
6 big bets or less in their stack, they haven't had any action
for 15 minutes because of the break, so it's shove and hope time
for them. Hey, that \Js\eigs\ looks pretty good.

If you've got a big stack after a break, you can nurse it if
your preflop cards are ordinary, or take some confrontations
where you'll be the favourite if your preflop cards are good.
Just after a break, the table will be very loose for a round as
the short stacked players take their chances.

