\chapter{Classifying your opponents}

% Last updated: 20200406

In this chapter we look at the four different types of players in
poker. Adjusting to each type of player is at the heart of the idea
that in Poker you play the player, not the cards. The kind of player
you're up against is one of the most important things to consider in
your poker decisions.

\section{Two dimensions, Four styles}

In poker you assess your opponents on two dimensions.
The first is called Tight/Loose and is a measure of the quality
of starting cards that a player will take a flop with. Someone
who folds nearly all their starting hands, and
only enters the pot with Queens, Kings, Aces, Ace King and
Ace Queen only when suited is ultratight. Someone who regularly plays
Jack-Three suited and Queen-Nine offsuit is loose.

The second dimension is called Passive/Aggressive and indicates
whether players bet and raise their hands, or are more likely
to check and call. How much betting heat do they normally generate
when they're in a hand? This dimension used to be called ``Weak/Strong''
but is now called Passive/Aggressive. ``Weak'' is now used in poker
writing to mean a scared, straightforward player, who can be
easily pushed out of pots when he doesn't flop top pair or a good
draw.

A Passive player is unlikely to put you under pressure with bluffs and
raises, unless he has the card strength to back it up. An aggressive
player is putting you to the tough decision with bets and bluffs and
raises; which is great when you've got the card strength to call him
and win big pots. A tight aggessive player is often making those bets and
bluffs and raises with a strong hand, so when you do call, you're
the one losing the big pots. Aggressive players put you to the tough
decisions.

Our two dimensions are independent of each other, like the X-axis
and the Y-axis in maths. You can be tight and passive, tight
and aggressive, loose but passive and loose and aggressive.

One part of life with two different dimensions is Intelligence and
Spirituality. Atheists in IT love to slander Christians as dumb,
but let me tell you there's plenty of dumb Atheists and plenty of
smart Christians; just taking dumb and smart to mean how good and
intelligent they are at IT. The spirituality dimension is a different
axis to the intelligence dimension, and someone's position on one
dimension doesn't tell you their position on the other dimension.

% FIXME: LaTeX Box magic

Four combinations of these qualities are possible. \\
Loose/Passive, Loose/Aggressive, Tight/Passive and Tight/Aggressive. \\
There are nicknames for each style of play: \\
Loose/Passive: ``Calling Station'' \\
Loose/Aggressive: ``Maniac'' \\
Tight/Passive: ``Rock'' \\
Tight/Aggressive: ``Expert''

There are three letter abbreviations for each style, LPP = Loose
Passive Player, LAP = Loose Aggressive player, TPP = Tight Passive
Player and TAP = Tight Aggressive Player. You'll see that in some the
literature and it's good to learn and use in your own notes. Nowadays
Loose Aggressive is called LAG, the G being the second letter in the
word AGgressive. So we have LPPs, LAGs, TAPs and TPPs.

% FIXME: Cite the Schoonmaker book properly

An entire book has been written on just this two-dimensional view
of player characterisation. It's called \textbf{The Psychology of Poker}
by Al Schoonmaker. It's a great book, though it concentrates
on Limit poker betting structure and has a lot of Seven Card Stud
content. Phil Hellmuth's \textbf{Play Poker like the Pros} has a
similar system, again focusing on Limit betting, though Holdem only.
Phil uses animals to characterise the player types, his animals are
as follows: \\
LPP=Loose/Passive: ``The Mouse'' \\
LAP=LAG=Loose/Aggressive: ``The Jackal'' \\
TPP=Tight/Passive: ``The Elephant'' \\
TAP=Tight/Aggressive: ``The Lion''

%% Phil can't help but add a fifth animal: ``The Eagle'', one of the
%% top 100 players in the world who occasionally comes down from
%% the sky and swoops on all your chips. You are unlikely to play
%% any of the top 100 players in the world in a Fast Pub game, they
%% have their own big games to play in. Although Jeff Lisandro
%% (tournament winning of around four million dollars, and
%% probably our biggest ever live cash player) played in
%% a free APL game in Sydney once, so anything's possible! I doubt
%% Jeff brought his best game to the APL! Many
%% of the bigger Sydney players will play in \$100 to \$500 buyin
%% games when they're offered, so keep an eye out for the experts
%% when you play in these games.

You can see from the nicknames and Phil's animals that the style
of play admired is Tight/Aggressive. it's the style of play you want
to be playing in all forms of poker, and the style you must play to
bring your game to higher
levels.\footnote{In recent times, the Loose Aggressive style
(``LAG'') is coming to the fore. When you feel more expert
at your game you can try broadening the hands you play, which
will make you harder to read, but start with fundamentally
tight aggressive play. Playing a wider range of hands means you'll
often be flopping middle pair not top pair and you'll be putting
yourself to a lot more tough poker decisions than you face as a tight
player.}

Hellmuth advises new players to start tight, only playing the
top 10 or 15 starting hands. He offers this excellent
advice to his readers: ``After all, how much fun can it be to
play so many pots and lose most of them''?

In Pub Poker there some other factors you might want to assess
players on, such as their board reading skills and their general
awareness of the game. ``Card sense'' I'll call it.  I'll cover more
dimensions at the end of this chapter but Loose/Tight and
Passive/Aggressive are by far the most important factors to
understand. For now we'll concentrate on those
two dimensions and how to use them.

Online poker players want to know the playing style of their
opponents, and they can't see the other players so there's no point
looking for Mike Caro tells). Automated stats programs tell online
players how many pots an opponent enters and whether he's limping into
them or raising. This gives a player a feel for whether an unseen
player is Tight or Loose (is he entering a high percentage of pots
preflop?) and whether he's Passive or Aggressive (is he raising the
pots which he's entering?). One online stat you may have heard of is
VPIP, which is the measure of how often a player voluntarily puts
money into the pot (when he's small blind or big blind it's not
voluntary). Someone with a high VPIP is loose, he's entering more pots
than he should, given the average distribution of cards. Someone with
a low VPIP is tight, entering fewer pots than you'd expect.

\section{Four player types}

Let's look at the four styles and how people in that style play poker.

\subsection{Loose/Passive: The Calling Station}

These players are the easiest to play against and win chips from.
You get a good hand and you put in value bet after value bet and
get paid off -- often by having top pair with the better kicker or
just top pair in the first place. You have KQ, the flop is Q84, and
you get called, called, called by Q5. The player's in the pot in the
first place with Q5 because he's loose, and letting you keep the
betting lead by just calling, calling, calling. At no stage in
the hand are you put under pressure with a raise. The calling
station's bad starting hand of Q5 is never strong enough to raise with
anyhow; he's just hoping to spike two pair against the odds.

Many complete beginners play poker this way. If someone makes
a few beginner's errors (such as never knowing when the action
is on him) and is staying in lots of pots, he's a calling station.

Very friendly, social people are often calling stations.
They feel as though raising is rude, and a good hand deserves
to win a few more chips, so folding seems rude also. Since
they're winning less with their monsters (by not raising)
and losing more with their losers (by calling) they are
Win Small/Lose Big players. They call too much, staying
in hands they have very little chance of winning. They fold
too little, especially preflop. They raise too little, so when
they do raise it's easy for their enemy to put them on a strong hand
and fold.

Calling stations are the Nice Guys of poker. The Old Poker Saying says
that Nice Guys finish last in poker. Now you know why.

\subsubsection{Beginner's Luck}

Calling Station's get a small amount of Beginner's Luck in poker.
When the Calling Station calls every street with his Q5 against my KQ
and the board is Q8425, he wins a pretty nice pot with his two pair
because that river sure looks like a brick to me. They don't get
enough of this luck to survive in the game and they get value bet to
death when they miss, which is most of the time.

\subsubsection{Playing against the Calling Station}

Calling stations can drive you mad when they river two pairs,
or hit some raggedy low flush with a garbage holding like 94 suited.
Because they're passive, they don't themselves bet or raise unless
they've got a pretty strong hand. Once you've picked them as
Passive, you can beware of their rare bets and raises, and get out of
their way unless you're very strong yourself.

Don't try and bluff calling stations. Go for the value bet,
and even the overbet. Charge them too much for their loose draws.
A calling station will occasionally beat you by making two pair with a
low river card that you're sure is a brick. That's the main element of
Beginner's Luck in poker; beginner's stay in with bad hands and make
weird two-pair hands on the river, against the odds, more than good
players.

Even when a loose holding does luck a good hand, it's often
not a strong enough hand to bet. 94 suited makes only a Nine-high
flush and loses to a higher flush. When 94 makes two pairs,
it's rarely top two pairs. On a board like T964Q the two pair 9944
loses to respectable two pairs like T9 and QT. So their
holding ends up making a poker hand that's ``not strong enough to bet,
good enough to call with, too good to fold'' which is a recipe for
small pot wins when ahead and large pot losses when behind.

\subsubsection{The Payoff Wizard}

A few calling stations pay off extremely big bets on the river; far
above the value of their own hand and the hand that the betting action
suggests the winner has. These players are called Payoff Wizards. If
you find one of these players at your table, make your value bets a
lot bigger against this player. Payoff Wizards are a gift from the
Poker gods; especially in the early rounds when stacks are still
deep. The really good Payoff Wizards bet tiny on the flop and the
turn, which lets you know they've got something themselves and also
lets you know they'll pay you off if you hit your hand. I've drawn to
gutshot straights plenty of times against Payoff Wizards, and it's a
beautiful thing when the draw gets there.

\subsection{Tight/Passive: The Rock}

After busting out in early rounds of their first half dozen
tournaments, people soon figure out that it's best to take flops
with good starting cards. They still don't bet or raise without
great cards, but they're now in the pot with better starting hands.

True rocks will only enter 10\% of pots. Apart from when they're
in the big blind in a limped pot, they're only taking one
flop per round, unless they hit a run of unusually good starting
cards. If the other players are bad and haven't noticed
he's a rock, he'll still get paid off on his ultra-rare bets.

Some of the ``make money playing poker online'' schemes use this
approach. Play ultratight, and get paid off when the good cards
come. In ``Rounders'', Matt Damon's voiceover suggests that
the target for a grinding Fixed Limit player is two big bets per
hour.

Rocks enter so few pots that they're hardly noticed
at the table. Many of them are quiet anyhow,
a persona which helps them look innocuous and get paid off by the
more aggressive players. You might hear comments from the table
when a rock finally enters a pot or (even rarer) raises. Recognize
the raise from a rock as what it is: an indication of a very very
strong hand.

Older players are likely to be rocks. I've also seen players who've
taken life-changing bad beats be rocks for months and years
afterwards. A life-changing bad beat is one that stops you getting to
the final table as massive chip leader in a big quarterly tournament
with first prize of \$10,000. I once made a guy fold KT on a KK54 flop
because he thought he was always getting bad beated again.

% FIXME: move this next stuff to the stereotypes on women section?
Some women are rocks, especially preflop. They
need a huge hand to raise preflop and just about Aces to
re-raise. I've seen (and paid off) many ladies who limp in late
position holding Ace-King suited. If they won't raise a pot with that,
a preflop raise from them pretty much means at least Queens. Be very
careful of a preflop raise from a tight woman. She isn't
loosening her play, she's actually holding a very
big pair. Fold your A9 or QJ as soon as etiquette will allow.

Some of the more aggressive guys jump into a contest with a
preflop raising woman, and pay off her very strong hand. I
used to do this myself, but I've grown out of the habit. This
isn't smart play at all. Try and be ego-less, and use your
superior player classification skills to get out of the way
of a preflop raise from a very passive player. Even the most
passive player in the tournament is dealt pocket Aces now and
again.

Rocks have the ``Lose small'' part of gambling mastered. Rock play
can be an effective style of playing cash poker. It's not a
good style for Fast Tournament Poker because Rocks don't win enough
on their few good hands and the fast rising blinds put them
under pressure. Female rocks do a bit better than Male rocks, because
some of the guys at the table don't bet hard against them and don't
push them off hands as readily as they do against the male rocks.

\subsubsection{Playing against the Rock}

Rocks can frustrate you because they stay at the table for a few rounds
and you can't get rid of them quickly because they play so few hands.
It takes some time for the blinds to eat them up (although it's
pretty quick in fast tournaments). Rocks avoid big confrontations
and tough decisions by simply being out of the hand preflop so often.

If maniacs try to break them quickly,
it's like they're smashing their own heads against a rock.

You want to blind steal a lot on the Rock's Small Blind and Big Blind.
If the Rock calls a decent flop bet or raises on any street, just get
out of her way. Against a maniac you might take your chances with
top pair decent kicker or even middle pair, but against a rock you
simply fold and take a small loss on this hand.

\subsection{Loose/Aggressive: The Maniac}

These are the most noticeable players at the table. They're constantly
drawing attention to themselves with their bets, raises, table chatter
and behaviour. If someone gets under your skin at the table, chances
are excellent that he's a Maniac.

Drunk players are invariably maniacs. They lack the patience to be
tight preflop and they also convince themselves midway through a
hand that winning \textbf{this} pot is the only thing that counts, so
they'll make a big bluff at it.

If a maniac gets good cards, look out! He can go on a rush like
nobody else. His aggression steals him pots, and when
he has the goods and someone takes a stand against him, his
stacks rockets up even faster.

Following my advice in this book, you'll be projecting a Loose/Aggressive
image. This is great both for stealing pots and for getting paid off.
I've had tournaments where for five hands in a row I've either folded
the table to my bluff, or had two stacks go all in smack into my nut
full house. The strategy in this book is to play Tight/Aggressive
poker, but have a Loose/Aggressive image.

The best movie example of a player pretending to be drunk but playing
smart is Paul Newman in ``The Sting''. In one of the best poker scenes
before ``Rounders'', he sets up Robert Shaw's Chicago mobster,
Doyle Lonnegan, in a game of five card draw played on a train.

Maniacs are good at the ``Win Big'' aspect of gambling. They splash
their chips around that everyone at the table realises that they
can't always have a monster hand, and their big hands often get
paid off and win big. Unless they've got great poker judgement, they're
also good at losing big. And because the game is No Limit, they can
lose a lot of their gains, or even their entire stack, in just a
single overplayed losing hand.

\subsubsection{Playing against the Maniac}

The first thing to realise is that you'll have to show down a winning
hand against the maniac. They're not easily bluffed. You want to set up a
trap against them, have a very strong hand yourself and let the maniac
bet into you the whole way. Don't raise or check raise until
the turn, by which time the maniac will probably be pot committed.

You don't win regular small pots against Maniacs, you look for a single big
pot, even an all-in, that will put you and the Maniac up against each
other with you having a big advantage. The maniac will sometimes bust
you by hitting a five outer on the river, that's part of poker.

When you've got a good starting hand it's a good idea to Isolate
re-raise the Maniac preflop so that it's just the two of you in the
hand. You don't want a third player sticking around and catching a big
flop when your aim for the hand is to get all the Maniac's chips.
You want to play this hand Heads Up against the Maniac, giving you an
exclusive shot at winning his chips. A double up against a Maniac is
much easier than being in a three-way pot and not knowing where you
stand against the third player.

\subsection{Tight/Aggressive: The Expert}

Tight/Aggressive is the style of poker the pros all move to. They're
in the pot with quality cards, and their bets and raises put the
other player to the tough decision, and give them information on
the other player's hand, helping define their own hand. They're great
at sensing weakness and being the player who bets all-in and gets
the weaker player to fold.

Expert players are also great blind stealers and card readers.
They have complete understanding on
the concepts of position, chips and bet sizing. They're working on
their game continuously, taking notes, staying alert and playing
their best game of poker.

To do well at pub poker, and to take your game to the next level,
you have to be aggressive in poker. Passive just won't get it done.

By folding more rather than calling, the Tight/Aggressive player
has the ``lose small'' quality of the Rock. And by raising more than
calling, he has the ``win big'' quality of the Maniac. Lose small, Win big.
It's what poker is all about.

Compared to the other play types, Tight Aggressive players fold more,
raise more and call less.

\subsubsection{Playing against the Expert}

The best way to play against experts
is to treat them with respect. There's plenty of other chips
from worse players to be mopped up before you have to run into
confrontations with the experts. Of course once you're at the final
table you'll have to confront the expert, and if you hold a nut hand
early on you'll extract chips from the expert. Outside of those
situations, stay out of the experts way and take the easier chips
on offer from the rocks, calling stations and maniacs.

One trick worth trying against the expert is an overbet. When you
hold a power hand on the turn or river, instead of betting half the pot
to the pot, try overbetting to one and a quarter times the pot. This can
throw the expert's hand reading off-kilter, he'd correctly fold to
the normal bet, but he might get overly suspicious of this strangely
sized bet and think that it's a big bluff and he'll pay you off.
The Expert's thought process runs something like this: ``Wouldn't he
be trying for a small surefire payoff if he had a strong hand?''.

You can also try ``honest tells'' against him: Strong means Strong and
Weak means Weak. Since there's not much trapping involved in fast
tournaments, honest betting and fast playing works best anyhow, but
if the expert is looking for Mike Caro tells and using
``Weak means Strong'' theory, you can try reversing that and maybe
the expert will hang himself on his own tell-searching.

\section{Players expect you to play like them}

``Players expect you to play like them''.\footnote{Except the Expert, who
has figured out your style of play, and has made the adjustments
needed to exploit it.} This is powerful advice.
Once you've got that figured you
can go a long way to manipulating other players into doing what
you want them to.

A rock gets pushed off hands because he expects other players to
have strong hands also. He raises preflop with pocket Jacks,
you call him from later position with 93. Flop is AT6. He checks, you
bet half the pot. He says ``I know you've got the Ace'' and he
folds the Jacks face up. You fold face down, say ``Good read''
and try not to smile as you pull in the pot.

If the rock check-raises you, Hollywood for three seconds then fold.
You're facing a strong move from an extremely passive player. It
doesn't get much stronger than this.

Maniacs expect other players to be maniacs. So they will call the
very rare bets from rocks and calling stations because ``he could
be bluffing here''.

Because a maniac is always suspicious of bluffs, he calls way too much.
This behaviour is at the heart of the poker saying ``You can't bluff
a bluffer''.

Some ``Experts'' figure out the Maths of drawing hands and conclude
that someone who called against the odds couldn't have made a longshot
draw such as a straight. These Experts are often overbetting top pair
good kicker and get upset when somebody draws out on them.

Calling stations call the bluffs of maniacs because they're calling
stations. A terrible mistake maniacs make again and again is to bluff
and bluff at calling stations. Remember the way to play them is to
value bet and overbet your made hands. Throwing big bluffs at a
calling station is a fast track to disaster.

\section{Some other dimensions}

While loose/tight and passive/aggressive are the two most important
factors to assess other players on, there are a few other things to
look out for. Don't strain your brain trying to pick up every little
thing, but here's a few extra factors to make this chapter complete.

\subsection{Card sense}

Does a player use poker terminology correctly? I hear some hilarious
mistakes at the tables: someone called a C bet a ``continuous bet''
and someone else told me once they had 20 outs with a low straight
flush draw. I love players who think they're giants of the game
and actually aren't. They're always good for a chuckle.

Extreme beginners will have no card sense at all, which will be
obvious in the first few hands when they don't know when it's
their turn to act, don't know about posting the blinds and show
similar inexperience with live poker. Some will call on a flop
of AQ4 rainbow with 87, holding no pair and no draw. These people
are welcome at any poker game in the world.

\subsection{Risk}

Is a player cautious or wild? Wild players take big risks and
enjoy the gamble in poker. They're happy to gamble all their chips
in one hand where they've got a decent draw. Some players happily
go against the odds or make crazy preflop plays with little upside,
such as shoving all in when deepstacked holding their favourite
hand, pocket fours.

Risky players will be winning and losing abnormally large pots
early in tournaments. Cautious players will win and lose steadily and
be more patient until their stacks are small relative to the blinds,
Which is a great way to begin a tournament.

\subsection{Trickiness}

Some poker players bet very straighforwardly. They'll put in a
Continuation Bet here and there, but by the turn and river their bets
mean pretty much what they say. A bet on the turn after a bet on the
flop means Top Pair Good Kicker or better, a Check on the Turn is
either shortening the hand by one betting round or giving up the lead,
and so forth.

Better players are changing their betting, sometimes checking when an
Ace hits the flop so they can get more value from the hand on the
turn; check-calling with sets; small bets first-to-act that could be
blocking bets, or sucker bets that actually want to be raised. A very
tricky player will try a raise-bluff on rare occasions; against most
players you should give a raiser a high degree of respect and put that
player on at least two pairs and often a set. Respect the raise.

It is possible to be too tricky in Holdem, and I warn in a few places
in this book not to slowplay. Pub players are more likely to call so
the main technique in pub poker is to get overpaid for your winning
hands; not to check and sucker a big bluff from an opponent. There's
normally not enough aggression at a pub poker table to get much
value from slowplay.

\subsection{Chip Shuffling}

Chip shuffling is the most useless trick known to poker players.
Someone who spends all of their time shuffling a few chips one-handed
isn't necessarily a good poker player. He just likes shuffling his
chips.

Personally I like to line up the side markings of the chips so there's
a good five and five appearance on my ten-high stacks. This is just as
useless a trick as chip shuffling, but it's more fun to do, for me.

\subsection{Dealing}

Someone who's not a good dealer is not a good poker player. Few
players do it to the level of expertise that I describe in my Dealing
chapter, but some people are just terrible dealers, especially in not
following the action and not knowing the correct time to produce the
turn card and the river card. Expect the bad dealers to be just as bad
as players.
