\chapter{Sport Structure, Poker Structure}

% Last updated: 20200425

Winning a poker tournament is like batting and scoring a hundred in cricket.
You need to build your innings, and not get out early
with a rash shot. The bowling and fielding might be good or it might
be easy. You might be hitting the ball well or be struggling and out
of form. Sometimes you'll get bowled by a swinging ball and
there's nothing you can do about it.

\section{Cricket Structures}

Test cricket and One Day cricket are different games and the ultra-fast
Twenty-20 cricket is another game altogether. Ball by ball the
action is much the same, but the different formats mean
different skills are required in each game.

Test matches are the pinnacle of the game. Test statistics and records
matter most to the fans and the players. There are some players who excelled
at Test cricket but not the faster paced One Day matches, while
other players (famously Michael Bevan) were One Day
specialists and never reached a position of comfort in the Test team.

Big buyin, deepstack slow poker is Poker's Test Cricket.
Small buyin, shortstack fast poker is Poker's One Day Cricket.
And ultra-fast tournaments, such as 30 big blind stacks with 10 minute
blind doubles, are poker's hit and giggle Twenty-20 games.

This book is all about how to play One Day Cricket Poker.

You won't play in a \$550 entry 50,000 starting stack 1 hour blind double
game very often, but you can play \$10 5,000 starting stack 15 minute
blind double games five times a week. The \$10 game is very
different to the \$550 game.

A good batsman might average 55 in Test Matches and 40 in One Day
Matches. The fast pace of the One Day game forces him to take
chances earlier and more often than in Tests. The important thing in
Test Matches is often staying in, whereas in one day matches
the focus is on getting runs quickly. There's an extra
measure of how quickly you score runs, Strike Rate, that's mentioned
a lot in short form cricket but not so much in Tests. If you get out,
you're leaving it to other batsmen to score runs, and there's another
one day game in three days time anyhow.

In a Test Match you can wait for the bad balls, or for the bowlers to
get tired. In big stack poker you can wait for good hands, simply folding the
mediocre starting cards and waiting for the better hands. In One Day games
and small stack poker you've got to take chances sooner; swinging at
good balls in cricket and taking chances with lesser hands in poker.

Any cricket fan can tell you the difference between the Test,
One-Day and Twenty-20 formats. But many pub poker players haven't given
much thought to the difference between deepstack and shortstack
tournaments. It's the same game they saw on World Poker Tour,
right?

In all forms of cricket a bowler runs in to a batsmen who
can leave the wide ball, play it and score runs from it, or get out
to it. In some games there's extra restrictions on where the
fielders can be placed and how many overs a bowler
can bowl, but the mechanics of each delivery in cricket is the same.
The mechanics of each hand in No Limit Holdem is the same.
But just as the cricket games are very different from each other, the
poker games are very different also.

There's no shame specialising in fast poker tournaments. Indeed,
the extra thought you put into structure and risk taking can give
you the advantage over more experienced players who are better in
slow deepstack tournaments, but don't understand fast tournaments.

\section{Tennis}

Tennis is another sport with differently paced competitions.
The prized tournaments are the four Majors (Australian Open,
French Open, Wimbledon and US Open) with a full field of 128 players
and men's matches best-of-five sets throughout. These slow tournaments
take two weeks to complete.

Smaller ATP tour events have a field of 32, 16 or 8 players in the draw
and early round matches (and sometimes even the final) are best-of-three
sets. There are a few of these events happening in different cities in
the world at the same time, players pick and choose which events
to play in.

The ultra fast hit-and-giggle Tennis format is called Fast-Four.

A few retired tennis stars have done very well in Poker. Russian
champion Yevgeny Kafelnikov has won a few big Poker tournaments.
Tennis legend Pete Sampras is a keen player. World Poker Tour
commentator Vince Van Patten was a US tennis star in the late
70s and early 80s, reaching 26th in the world in the heady days
of Vitas Geraulitis, John McEnroe and Jimmy Connors.

Some poker wags have named the hole cards Ace-King Anna Kournikova, in
honour of the Russian Tennis beauty with the initials A.K. Like
Kournikova, grumble the poker players, Ace-King looks good but never
wins.

\section{Golf}

Similar to Tennis, Golf has four Major tournaments each
year---the British Open, US Masters, US PGA and US Open. Smaller
tournaments have smaller fields and some are played over
Three rounds (54 holes) not Four rounds (72 holes).

There's actually a second string golf tour where players pay an
entry fee and compete for each other's money, just like
Tourament Poker players do.

Golf, Tennis and Poker are individual sports, where
the normal competition format is ``every man for himself'' and
you have to come out on top of the field to win the tournament.

In a knockout Tennis tournament, you get helped
if star players lose to lesser players and you get to play the
lesser players in later rounds. In Poker tournaments you're
helped if star players lose to lesser players, often on other
tables. In both sports, you can only beat the players you're
up against on the day.

In a Tennis Tournament, the matches get harder as the tournament
progresses, because of the seeding system and because in the Third
Round you're playing someone who has also won their first two games
and in  the Fourth Round you're playing someone who has also won their
first three games. The penalty for losing a match gets greater as the
tournament progresses; the prizepool jump gets bigger with every match
you win.

In Golf, you get partnered with better players each round, and the
lowest scorers after the Third Round are the last playing group on the
Fourth and final Round. Although your final score is a cumulative
total of all the strokes you've had over the four days, the penalty
for missing a short putt on the final hole of the final round is a lot
more than missing the same putt on the sixth hole of the second day.
In the same way, a mistake on the final table has a magnified impact
on your tournament success compared to the same mistake before you're
in the money.

% \section{What makes a champion?}

I've read my fair share of sports biographies and many people read
these books looking for inspiration in their lives. Successful
sportspeople and poker players are hard workers, competitive and
persistent. They overcome adversity, and don't listen to the
people in their lives who tell them they won't make it.
