\chapter{Freeroll Win}

% Last updated: 20181129

While I've been working on this book I've been playing more
pub poker and I won a 41 player Freeroll at my local
Bowling Club on Wednesday 7th November. Here's how I
did it.

\section{Game conditions}

This game had almost all of the Niggles that I hate.
Spend cards, early bird bonus, and a token extra \$2 fee
for double stack start, on a Freeroll. Three of the players
were randomly nominated as Jackpot Bounty players; if they won the
tournament tonight they'd be given an extra prize.
I mentioned that Niggles are more prevalent in the Freeroll world, this
game lived up to my suspicion.

Although the Spend Card had four squares to fill in, all you
needed to do was spend \$10 at the bar or bistro and bring your
receipt (it was too much work for the staff to stamp the actual
cards but they'd happily print you a receipt which they'll do
for any transaction at any time anyhow). So buying two beers
before the game sorted out my Spend Card needs.

I don't as a rule normally play Freerolls but I'd been stuck
at home all day on a Wednesday and was feeling a bit Stir Crazy.
A 5 minute drive up the road to the Bowling Club seemed like
a good idea. If nothing else, it would give me some real life
experience of Freerolls to write about in this book.

\section{The Structure}

The game had a 4,000 start stack, but with the "double stack for \$2"
and extra 5,000 for spend card I started with a huge 13,000. Most of
the expected early blind levels were there. 25/50, 50/100, 100/200, 200/400,
300/600. Once the blind hit 500/1000 it was just about full doubles
from there: 500/1000, 1000/2000, 2000/4000, 5000/10,000, 10,000/20,000
and 20,000/40,000. Once the 5000/10,000 limit was reached, the TD cashed out
all the 1,000 chips, which made for little betting subtletly, you could
no longer raise to 12,000 because you only had 5,000 chips left. The game
became the ultimate White Chip Poker.

\section{Early stages}

I was at an 8-seater table and folded a few hands before
making my first position play of the night, getting someone
to lay down a pretty good middle pair hand when I had nothing.
Because I'd done the spend card and paid the niggly \$2 I actually
had the chip advantage over many of the players at the table
(it's actually nice to be on the \textbf{good} side of an
uneven playing field!).

In the first level I did a pot-builder preflop raise with T7,
got a bunch of callers and got the dream flop of 986! I cbetted
500 into an 800 pot and the player to my immediate left
insta-shoved for 5500 with Q9!! A third player (holding 87) tank-folded
and I had to wait patiently for my turn to call holding the nuts.
The runout was 986TA so I had to dodge a Jack on the river,
and just as well that the tank folder did fold his 87 otherwise we
would have split with me.

I made a couple of mistakes in the early level. On the button I held
22, put in a pot-build raise and got a dry flop of Q82, my cbet
folded three players and I won a small pot with a set, never a
good outcome. I should have let players see a free turn there, Q82
is a desert-dry flop.

Another hand I was underpaid for saw me with 95 on a QJ8 flop. I was
check-calling on the flop and turn and the runout was an interesting
QJ876; making me the second-nut low straight but really it
should be a winner, I doubt anyone had T9 for the flopped nut straight.
I bet out on the river and got the bettor and others to fold, I should
have check-raised on the river, the original bettor (holding JT)
would have value bet a third time but my strong lead got him to fold
as he didn't even have a two pair hand.

\section{Middle stages}

At the break I had a stack of 31,550 which I'll call 31k for simplicity.
This was after the 25/50, 50/100, 100/200, 150/300 and 200/400 levels.
With the blinds increasing to 300/600, I still had a 50 BB stack.

I went card dead for a while, but still making position plays here
and there. One hand I had As4s, flopped a nut flush draw but slowed
right down and missed on the river; I'd paid 5000 to see the turn,
got a free river after the board paired but missed everything and
lost to trips, good play by me to take the free river as I would
have been wasting chips with a turn bet and could even have been
check-raised.

After the blinds hit 500/1000, all increases were straight blind doubles.
The blinds started to bite and I hit some good hands, an under-the-gun
pocket KK limp-raising a lot and busting a smaller stack's 77. I love
the UTG limp-raise, and it was a good one here as on the runout of
AJ4J3 I could well have lost to an Ace or a Jack.

I butted heads a few times against the oldest player at the table,
who was quite a canny player. He was a rock preflop, never raising,
often folding, but limping with very strong hands like Ace-Queen.
He often flopped top/top and was able to call my bluffs on every
street and win plenty of chips from me. He did shove all in for
about 40,000 with blinds at 300/600 holding AA, picking up just
the blinds for that hand, which I thought was very bad play.

I doubled through him when my stack was down to about 25k; holding
Q8 I got the priceless flop QQ8. I called his bets all the way
until on the final board of QQ866 I raised his 2000 bet all in
for another 19k; he held 86 and it was too hard a fold for him
so he made the call.

The player on my left was shoving all in for his 25k after some
of my limps, I got tired of this and took a stand with 76. He
held AJ but I rivered a pair on a nothing board to send him
home. Some very good luck for me there, really I had superb
card luck the whole night. Soon I was at the final table of
Nine players.

\section{Final Table}

One of the busted players dealt for the final table and after
a slow start soon got the hang of it. My stack was at least
200k with about 570k in play. I won an early race 55 versus AK
on an all-low runout. A later race of 77 against AJ saw
both a Jack and a Seven on the flop and more chips to my stack.

I had enough of a stack to take my chances with QT suited on a
four way all-in against some small 15k-30k stacks. K9 on the flop
followed by a Jack on the turn saw me bust three players in
a single hand and immediately make the money in the last three.

It seemed that blind levels had been shortened to 10 minutes in
the late stages; a few people murmured that the tournament
always finished at 1030pm. I'll check this next week. Normally
I'd be livid at this disrespect of the players but it's
a freeroll and they can do what they want in a freeroll.

In my winning hand I had 98 to a flop of Q88 two diamonds.
I checked and instacalled an allin, the all-in player holding
KJ of diamonds to have flush outs, but none came in the runout
and I was that night's champion.

The \$350 payout of \$200/\$100/\$50 was polluted by
pub money and tickets into other games that I now feel
I'll have to go and play in. I'd have happily taken
\$200 in cash if offered, but, again, \textbf{it's a freeroll}.
Freeroll players have no right to complain about the prize pool,
or even capricious shortened final blind levels, since the
venue is putting up the money.

\section{Lots of luck}

I had lots of card luck in this tournament; early on T7 got a 986 flop,
in the middle stages my Q8 got a QQ8 flop and a runout of QQ866 against
86, just about the only runout where someone was going to pay me off big.
And my QT at the final table busted three players by turning a straight.
Just about the only bad card luck I had all night was when a flopped
Nut flush draw didn't get there. I'd rate my luck in this tournament
as 9.5/10.

My play was pretty good, I made a few mistakes against the tight old
player and shouldn't have given him so many chips. I'll mark myself
down on that score. But my position shot-taking was excellent, my blind
stealing was successful and I felt I had the run of the table and
the respect of the table from the get-go. 7/10.

I believe out of the 41 players in this freeroll, I was one of
the top five. The player who came second was a good player, the player
I busted with 76 also earned my respect and there would have been
some good players at the other tables who I didn't get to play
against. So I personally played against two other good players
and there would have been two others. So five good players,
including myself, in the tournament.

The combination of being one of the good players together
with my great card luck got me the win.

\section{Bad play in other hands}

There was some very bad play I got to see in hands I wasn't involved in.
The old guy I'd given a lot of chips to and who was a big chip leader
at my table at the time shoved all in for 19,000 with blinds at 300/600.
Understandably everyone folded and he proudly showed pocket Aces! This
isn't a great way to play Aces and you should be letting some people
take a flop and using your judgement to see if your Aces are still ahead.

In an earlier hand, the big blind check/called every street on a runout
of K33J8. The bettor had AK, the caller had 53. The caller's mistake was
not putting in a raise on any street; not even on the river where it
was obvious that if he had a 3 he was still good.
