\chapter{Game selection}

% Last updated: 20181018

When I first played poker there were only two places to
play, the casino and home games.
I used to laugh at the sections in US poker books talking about 
game selection, advising you to choose the games suitable for you at
the amounts you could afford to play. There wasn't that
choice available to me.

Today, players are spoilt for choice. On any night
of the week there's a game running in each of 
the major suburbs of Sydney, and three or four in
the city centre. Some venues run Poker Tournaments five, six or seven
days a week. You should have no trouble finding a game
you can enjoy and do well in. Here are some commonsense
suggestions.

\begin{itemize}

\item Play at a venue that's easy to get to and easy to go home
from. Make sure you're not under time pressure to finish a game.
Always have good safe transport home afterwards.

\item Play at the stakes you're comfortable at. A \$10 
game with a \$200 first prize might be peanuts to you. 
A \$100 game with a \$2,500 first prize could be more your thing.

It should hurt a bit when you lose a Poker Tournament; not 
throw-yourself-off-a-bridge hurt but you don't want to play for
Matchsticks either. Everyone's financial position is different,
find the level that you're happy in.

If you're a student, \$100 games might be something
you only play once a month, and a \$10 game twice a week is
a better use of your money and time.

\item Don't play every night. I find that when I play
three nights in a row I hate poker on the third
night. There's some players I see every time I go into a venue,
I don't think they have as much fun with their poker as I do.
There's much more to your life than poker, sometimes the smart
move is not to play.

\item Play where you're interested in the other services
that the venue has to offer. If you like the food or drink or
TAB or sports coverage or other facilities at a Club you can 
have a good time there even if you're out of the poker early.

I encourage Pub Poker venues to show
AFL games live. The NRL often has three-in-a-row Super Saturdays
where you can only watch the games on Pay TV; playing a Saturday
Poker Tournament while watching the games is a great option
for League Fans. I like well maintained
Pinball Machines. I also love full sized Snooker Tables with good
equipment. A beer selection that's wider than
just Tooheys, Hahn, Carlton and VB is great. Little Creatures from Western
Australia is a current favourite. Coopers beer, either bottled
or on tap, is always good.

I rarely eat at venues before Fast Poker Tournaments.

% The most exotic beer I've drunk while playing Pub Poker is an Austrian
% beer called G\"{o}sser. I had it a shooting club in Luddenham, in South West
% Sydney. I didn't play well at all that day, but the beer was fantastic.

\item Look at the bonuses available for that game. Is there a
guaranteed prize pool? Free food for the players? Sometimes
there's an overlay--- you might have 20 people paying \$10 each
to enter with a guaranteed prize pool of \$300. One venue kicks
in \$100 per 10 paying players, so if 34 people play that night
there's \$340 + \$300 = \$640 in the pot, which is a nice incentive.
I'm always happy to play in that game.

\item Bring some friends along or play somewhere that you'll
know a few people. Poker is a lonely game at the best
of times and it's good to have some people even just to share
a bad beat story with. Though I also like
to go to a new venue alone and face the challenge of
reading entirely new players, and being an unknown player.
I'm always on the lookout for new games and new venues that
might be to my advantage.

\item Play somewhere that you respect the tournament directors
and the poker room setup. If the equipment (tables, chairs, chips,
monitors) is excellent you know the venue has made a commitment
to running some good poker. Lighting and air conditioning
play a large part in your comfort at the table.

\item If a venue close to where you live starts
up a poker game, go along and support it. Even if it turns out to
be a bad game or folds after a few weeks, you've met some people
from around your area interested in poker, which might introduce
you to other games around the area or some home games.

\end{itemize}

In the chapter ``The Pub Poker Price'' I write about some
aspects of Pub Poker that I'm not so thrilled at.
