\chapter{Made Hand, Drawing Hand}

% proofread: 20250430
% CHAPTERLENGTH: 2

Old poker saying: ``poker is a battle between a made hand and a
drawing hand''.

Hands like top pair/top kicker, two pairs and sets
are usually the best five card poker hand anyone can
make after the flop. But by the time that the turn and river come out
and there's the finalboard of five community cards for players to use,
straights and flushes can be possible.

For the purposes of this chapter, I'm going to label some hands on the
poker ladder; ignoring the top hand (straight flush, very rare) and
the bottom hand (high card). Quads and full houses are \textbf{Monster
Hands}, flushes and straights are \textbf{Power Hands} and three of a
kind, two pairs and pairs are \textbf{Made Hands}.

A \textbf{Drawing Hand} is a four-card hand that becomes a power hand
when the right fifth card comes on the turn or the river. A drawing hand
is four-to-a-flush or four-to-a-straight.

A \textbf{Complete Hand} is a five-card poker hand where
you're using all of your five cards to make the hand, there are no
kickers. The complete hands are full houses, flushes and
Straights. Quads is by this definition not a complete hand
but if you make quads using one of your hole cards you're
unbeatable so we'll treat it as a complete hand also.

% FIXME: do this in a table?

In a made hand versus drawing hand matchup, the made hand bets on the
flop and the turn, and the drawing hand calls trying to hit a
complete hand. Most of the time the made hand is a big favourite;
often a 65\% favourite on the flop and a 80\% favourite on the
turn. I'll look at the bet sizing implications of these numbers.

\section{Wet flops, Dry flops}

When the flop comes, have a look at how draw-heavy it is. The flop 
\Ks\eigh\trec\ has no straight or flush draws possible. This flop is
called a \textbf{dry} flop. The flop \Jd\tend\eigs\ supports a diamond
flush draw and someone could already have a straight. This is a
\textbf{wet} flop. Some flops are so wet that it's possible for
someone to already have a power hand of a flush or a straight. On most
flops a flush draw or an open-ended straight draw is possible.

% FIXME: How many flops are rainbow? How many flops are widely spaced?

As the turn and the river are added to the board, the drawing hand
either gets there, turning into a power hand, or it misses and ends up
worthless. The player who flopped the made hand will stay in the lead
when the drawing hand misses and win the pot. The made hand player
loses the pot (and probably more chips on the riverstreet)
when the drawing hand gets there.

The trick in no limit poker, when you're the player with the made hand
on the flop, is to charge the drawing player an amount that he will
call, but that is still a losing call for him. If
you do that properly, the calling player is making a mistake,
\textbf{whether he makes his draw or not.} Because of implied odds,
the amount of money that a drawing player can win isn't just the pot
he has to call his bet into; he could win more chips on the
riverstreet.

One challenge for the made hand bettor is that unless he has the best
possible made hand of top set, he doesn't know for sure that his
enemy has a drawing hand or a better made hand or a worse made
hand. Poker isn't always made hand versus drawing hand; sometimes it's
made hand versus better made hand; in which case the pot can get very
big very quickly. The most explosive confrontation like this is set
over set. A made hand bettor is trying to find a middle ground in
charging drawing hands too much to try to hit their draws, while also
trying to discover if the enemy might even have a made hand himself.

The highest of the flopped made hands (Sets and Two Pair) are themselves
drawing to the monster hands of quads and full houses. A flopped set
becomes a monster about 30\% of the time, flopped Two Pair becomes a
monster about 16\% of the time.

\subsection{Two-Way Hands}

A few holdem holdings flop both made hand and drawing hand aspects.
These are called two-way hands; they have two things going for them.
They've got a decent made hand, which could be best hand right now,
and they also have the chance to make a power Hand.

The best of the two-way hands is pair with flush draw, such as
\Kd\tend\ on the flop \Kc\nined\eigd\ . This holding has a flush
draw and already has top pair/decent kicker as a made hand.
Flush draws with an overcard can improve to a made hand by
pairing the overcard or a power hand by making the flush. \Ad\sevd\ on
\Kc\nined\eigd\ can improve to a pair of Aces with an Ace on the turn
or the river.

Two overcard straight draws can likewise improve to top pair or the
straight. \Jh\tenc\ on a flop of \eigh\sevc\tres\ improves to a monster
nut straight with a nine or a made top pair with a jack or a ten.

Next in the two-way Hands is pair with straight draw like
\tend\nined\ on the flop \tenc\eigh\sevs\ . This hand
isn't as good as pair with flush draw; because if the turn improves
the hand to two pair it can improves the enemy to a straight.

The last of the two-way hands is overpair with straight
draw. \Js\Jc\ on a flop of \tenc\nineh\eigs\ can beat any top pair
and has flopped an Open-Ended straight draw and should be the winner
if the runout brings a queen or a nine. Like pair with straight draw,
this hand doesn't want to improve as a made hand. If you improve to
trip jacks, someone could well have a straight on a perfectly
connected turnboard of \tenc\nineh\eigs\Jh\ .

% FIXME put this back? Commented out for now
%% Omaha, where each player gets Four hole cards, is full of two-way
%% and even three-way or four-way hands. KKJT on a KQ9 flop has top
%% set and top straight and King-high flush draw.

%% \section{High Card Holdem}

%% Late in Tournaments, as the blinds get very big and stack sizes diminish
%% to 15 BB or so, High Cards predominate. Small suited cards and small
%% connectors just don't hit often enough to justify calling bets with
%% preflop. They can occasionally be used as re-steal hands, but cracking
%% hands go down a lot in Final Table poker. Confrontations at the Final
%% Table are normally Made Hand against Made Hand, and whoever's flopped
%% the best pair wins.

%% \section{Exercises}

%% % Notice I get the chapter number from a LaTeX variable, not hard coded.

%% \arabic{chapter}.1 (YouTube) In the last two tables in the 2015 WSOP
%% Main Event, Joe McKeehen busts Justin Schwartz on a
%% \sixh\tred\twod\ flop, all the money going in on the flop. What hands
%% did McKeehen and Schwartz have?
