\chapter{A Quick History of Major Poker Games}

In the last 100 years there have been five major Poker games. We're
not interested in the really early days of Poker; whether there was a
20-card-deck French game called Poque or what was played by cheats on
the Mississippi Steamers in the days of Mark Twain. The poker games
that interest us are: Five Card Draw, Five Card Stud, Seven Card Stud,
Holdem and Omaha.

% FIXME: sentence like ``For each game, we will look at the amount
% of information available to other players.

\section{Five Card Draw Poker}

In Draw poker games, players have five cards to themselves and they're
all private. Here's a player's hand after the final draw in Draw
Poker: \back\back\back\back\back\ . If he drew one card in his draw,
he might've had two pair looking for a full house, four cards to a
flush, four cards to a straight, trips and he's only drawing one card
not two for deception, any kind of hand really.

You don't get to see any of the other cards in Draw Poker, so there's
not much information to go on.

Our main interest in Draw Poker today is in the odds of Five Card
hands. Holdem's community cards are a Five Card random hand, so poker
odds apply to the community (how often it will be a flush, how often
it will contain a pair and so on).

Movie: \textbf{The Sting} starring Paul Newman and Robert Redford.

\section{Five Card Stud}

In Five Card Stud, each player starts with a downcard and an upcard,
and after each betting round gets a new upcard, finishing with one
down, four up. Here's a heads up confrontation in Five Card Stud:
\back\nines\Qc\Jc\tred\ against \back\fourh\Jh\tend\trec\ . With
only five cards per player, the poker hands made are pretty boring;
high cards and single pairs are typically enough to win.

Five Card Stud was the main game played in the high stakes months-long
Heads Up confrontation between original WSOP Champion Johnny Moss and
Nick ``The Greek'' Dandolos. 

Movie: \textbf{The Cincinnatti Kid} starring Steve MacQueen and Edward G
Robinson. An extras feature on my DVD is a voiceover commentary from
Phil Gordon.

\section{Seven Card Stud}

Seven Card Stud is an extension of Five Card Stud. Each player starts
with two downcards and an upcard. As the hand progresses players get
three more upcards then a final downcard. A final confrontation in
Seven Card Stud is \back\back\nines\Qd\Jd\tred\back\ against
\back\back\fourh\Jh\tend\trec\back\ . Seven Card Stud is a nice mix of
visible upcards and mysterious down cards. It's a great game and I
wish it was played a lot more in Australia.

There was very briefly a game called Six Card Stud but it died out
very quickly as Seven Card Stud grew in popularity.

Movie: \textbf{Rounders} starring Matt Damon and Edward Norton.
What? Rounders is a Seven Card Stud movie?? It sure is. Apart from the
No Limit Holdem hands against Teddy KGB, and the Fixed Limit Holdem
hands at Atlantic City, all the hands you see in Rounders are Seven
Card Stud Hands. Mike and Worm beat the Trust Fund babies at Stud,
Worm cheats at Stud in Petra's club, Mike calls the hands at Stud in
the Judge's game and Mike and Worm get beaten up by the Police when
cheating at Stud.

\section{Texas Holdem}

Holdem is a variation of Seven Card Stud. Instead of everyone
having Four unique upcards, Holdem uses Five shared upcards. You don't
have to remember the upcards that other people folded in Holdem; the
shared upcards are visible throughout the hand. And with only two
private cards each and five shared cards, you make the Nuts quite often
and you know that you've made the nuts.

A final Holdem confrontation is
\back\back\nines\Qd\Jd\tres\fourc\ against
\back\back\nines\Qd\Jd\tres\fourc\ . Players share the same five
upcards.

\subsection{You know what the Nuts is}

The great attraction of the information you have in Holdem is that you
can see what the Nuts is. If the Board doesn't have a pair, nobody can
have a full house. For someone to have a flush, the Board must have at
least Three of a suit, and a flush is only possible in one suit. For
someone to have a straight, the Board must have three cards in a
continuum of five. Holdem Players soon learn to read the Board and to
recognize what the nuts is at each phase.

A Holdem player whose whole cards are \Ac\trec\ knows he's the winner
in the showdown \back\back\nines\Qd\Jd\tres\fourc\ . A stud player,
looking at another player showing
\back\back\nines\Qd\Jd\tres\back\ doesn't know for certain what the
other player has; he could have a flush in diamonds, a flush in spades, a
straight, a full house or even Quads. Being able to see what the Nuts
is is one of the great features of Texas Holdem.

\section{Omaha}

I won't say too much about Omaha or I'll never stop, but it's the
fifth major poker game. It uses the same Community as
Holdem, with the same Flop, Turn and River. Each player gets Four hole
cards and must use Two and exactly Two of his hole cards in the
Showdown (you can't play the board in Omaha). Our Omaha confrontation
is 
\back\back\back\back\nines\Qd\Jd\tres\fourc\ against
\back\back\back\back\nines\Qd\Jd\tres\fourc\ . Players share the same five
upcards. Because players now have nine cards to make a five card poker
hand from, you get some very high poker hands being made in Omaha.





