\chapter{Two Card and One Card Hands}

% Last updated: 20200519
% FIXME: git rename this file to twocard_onecardhands.tex ??

In Holdem you can use both, one or none of your hole cards to make your
best Five Card hand. If you use none of your hole cards you're
\textbf{Playing the Board}, and the best you can do is split the
pot, because other players in the pot can match your best hand by also
Playing the Board, or can beat it if they can by using a better hole
card of their own. On a final board of \Ad\nined\eigd\Jd\sixd\ your
\Qc\Qs\ has made a Diamond flush, but so has everyone else in the hand,
and if anybody has a single diamond higher than the \sixd\ they can
beat the board and beat you.

There's a story that in an early hand of the 2005 WSOP Main Event,
the final board at one table was AKQJT rainbow (no flushes
possible). Nobody can beat that board and anyone who stayed in would
split the pot. On the Finalstreet, one guy bet and a second guy
raised. Two inexperienced Internet guys folded, and the pot was split
two ways not four ways! I'm not sure if this is a true story or Urban
Legend, but it sounds one of those stories that are too good not to be
true.

\section{Two Card Hands}

A hand where you're using both of your hole cards, in a meaningful
way, to make your best poker hand is a \textbf{Two Card
Hand}. A Two Card Hand is stronger, and more deceptive, than a
One Card hand. Two Card straights and Two Card flushes are
strong, Two pairs and Sets are strong.

Two Card Power Hands are only strong on a Finalboard where no
One Card Power Hands are possible. If you've got AK and the top
straight on QJ84T that's great but if four of the Boardcards are
diamonds you lose to a One Card Flush. If the board has three
diamonds you can still lose, but now you can only lose to a Two
Card Flush; so your straight has a lot more value.

If you've got a high pair weak kicker, your kicker might technically
be part of your best five card but you're not using it in a meaningful
way and really you've got a Single Card weak Made hand. With K6 on a
KT942 Finalboard your six isn't really playing and you're
basically playing just your King for top pair no kicker. Anyone else
playing a King will have you outkicked.

\section{One Card Hands}

A One Card Hand is one where you're only using one of your hole
cards to make your best poker hand.

One Card hands of top pair are pretty weak; you need your second card
kicker to play and be good to have a confident made hand. A5 on a
board of AK9T6 is a single-card top pair; sometimes you'll win but you
lose to anyone else with an Ace unless they've got the even worse A4,
A3 or A2 in which case it's a split.

If you can make trips with a One Card Hand you're in much more
powerful shape. 83 on board of K88T4 feels like the winner. There's no
straights possible on that board, there's only one more 8 in play and
it's probably still in the deck not in someone else's hand.

\subsection{One Card Straights}

If the straight is made by a gap card in the board and you've got that
gap card, you should be the winner and only lose to an unlikely Two
Card Straight. On 9873J, your AT should win you the pot, but if
someone has QT you're toast. Someone with the Double Card straight 65
has been outdrawn; remember a Double Card hand loses its Power when
there's a Single Card hand possible.

When the board shows four-to-a-flush or four-to-a-straight, a One
Card Power hand is possible. It's obvious to everyone in the hand that
the highest single hole card of the flush-suit will win, or someone
with a card of the right straight rank will win.

When the board is this wet, your Two Card hand loses its power and
you've often been outdrawn. Your Two Card hand is only powerful on
boards where a Single Card Hand isn't possible.

A One Card hand is only possible after the turn or the river. If a
One Card hand is present on the turn, the flop must have been wet
(you don't get any one card hands by the turn after a flop of K93
rainbow). Almost any flop allows a One Card hand by the river,
given the right runout (K93 can runout K93QT making a One Card
straight for the Jack).

\section{Exercises}

\arabic{chapter}.1 AKQJT rainbow is by far the most common unbeatable
board, where everyone still in the hand splits the pot. AKQJT Royal
Flush is also unbeatable. Are there any other unbeatable boards?
