\chapter{Style Guide}

% proofread 20250418
% CHAPTERLENGTH: 2

\section*{Eight Player Tables}

Hands in this book assume an eight-player table. A lot of Pub
Poker tournaments use round eight-player tables with foldup legs and a
drink holder in the table for each seat. These tables are easy for a
tournament director to fold into a large disc and roll away as the
tournament field shrinks and tables are combined. A final table of
nine players is assumed.

k is used for ``1000'' in stack sizes; I had a stack of 27k and the
blinds were 1k/2k. When I do write out the full numbers I use
placeholder commas to make the numbers easier to read; I had a stack
of 27,000 with the blinds at 1,000/2,000.

\section*{Hole cards}

Offsuit hole cards use hearts/clubs, pocket pairs use spades/clubs and
suited cards use diamonds. This makes it easy for you to read the hands.

I shoved with \Ah\tenc\ but had no chance against \Ad\Kd\ and \Js\Jc\ .

In quick discussions I don't use card picutres I simply use
letters. The cards from highest to lowest rank are AKQJT98765432 and s
means ``suited'' and o means ``offsuit''. x by itself means ``any
card'' so Axs is a suited Ace and Ax is Ace-any card. You'll have no
trouble reading either format.

I shoved with ATo but had no chance against AKs and JJ.

%% \section*{Stack sizes}

%% Stack sizes are often expressed as a multiple of the Big Bet/Big Blind
%% (BB), rounded intelligently. With a stack of 26,500 with blinds at
%% 1,000/2,000 then you've got 13BB. A 39,200 stack at the same blind
%% level is 40BB. A lowercase k is used as an abbreviation for thousand,
%% the two stacks shown are 26.5k and 39k stacks with blinds at 1k/2k.

\section*{High to low flop}

The flop is always shown from high card to low card, reading left to
right. The flop is a simultaneous three card reveal, regardless of the
position of the cards you should think of a flop of 6K8 as being
``King Eight Six''. The turn is shown to the right of the flop and the
river is shown to the right of the turn.

The Board for Joe Hachem's 2005 WSOP Final Hand was
\sixh\fived\fourd\As\fourc\ .

\section*{Example hands}

Most of the hands in this book are from my own play, or hands I've
seen with my own eyes. A few hands are from the WSOP Main Event
of 2005, which was a memorable poker tournament for many reasons.

\begin{itemize}
  \item It was won by Australia's Joe Hachem. This created a lot of
    media interest in Australia which led to the Pub Poker boom.
  \item The ESPN coverage of the event, with commentary from
    Lon McEachern and Norman Chad, is extensive and
    excellent. Excerpts from this coverage can be easily found on
    YouTube. I own about five hours of 2005 Main Event coverage on
    DVD, which I watched in my early poker years, and I remember a lot
    of the featured hands very well. The coverage covers all stages of
    the tournament, whereas other poker shows such as World Poker Tour
    only show final table play. In my opinion, you learn a lot more
    about poker by watching WSOP hands than you do watching WPT hands.
  \item Dan Harrington uses four examples from this tournament in his
    book Harrington on Holdem Volume 3: The Workbook. If the hands are
    good enough for Dan, they're good enough for me.
  \item The start stack of the tournament was 10,000 in
    chips. Although blind levels were very long (two hours)
    players quickly became short on chips once they lost some pots.
  \item There were a lot of inexperienced Internet qualifying players
    in the tournament, chasing their poker dream after the wins of
    Internet qualifiers \textbf{Greg Raymer} in 2004 and \textbf{Chris
    Moneymaker} in 2003. In 2003, Chris Moneymaker
    borrowed some cash for his Las Vegas flight and expenses from his
    friend \textbf{David Gamble.} Both Chris Moneymaker and David Gamble are
    their real names. You couldn't make this up.
  \item The standard of play in the 2005 WSOP was roughly equal to the
    standard of play in a Sydney Pub tournament today. 2005 was before
    the era of Game Theory Optimal (GTO) solvers so the playing
    standard was roughly what you'd expect from good players before
    new tools in Online poker changed the game.

\end{itemize}

WSOP example hands are \textbf{At the WSOP}, hands from my play are
\textbf{At the tables}.

%% I've found I learn more from watching WSOP hands than I do
%% from World Poker Tour hands. I have WPT Season 1 on DVD and don't feel
%% a strong urge to re-watch it. World Poker Tour normally films just the
%% final table so you're already watching shove and call final table
%% poker. You can learn a lot of Final Table play from WPT, but for early
%% stage and middle stage play, watching WSOP hands will teach you more.

\section*{Language}

Male pronouns are used for poker players in most places in this
book. Over 80\% of poker players are men so it's not lazy writing
to say ``he bet 600'' rather than ``the player bet 600''. There's an
extensive \textbf{glossary} at the end of the book, that I spent a lot
of time on. Some of the words are of my own invention. Make sure you
understand the terms Flopstreet, Turnboard, Turnstreet, Finalboard and
Finalstreet. A person who puts money into the pot is the ``bettor''
and reraising preflop with Kings is ``better'' than flat-calling.
