\chapter{Themes}

% Last updated: 20200507

There are several themes that run through this book and are mentioned
several times. These are fundamental to good poker. Read the list now
and come back to this page often. If you go against
these principles, you're probably not playing good poker.

\begin{itemize}
  \item Win big, lose small
  \item Play tight/aggressive poker
  \item Make position plays
  \item Look for tells and follow the action
  \item Size your bets well
  \item Don't slowplay
  \item Raise to find out where you're at
  \item Respect the raise
  \item Know the structure
  \item Enjoy your poker
  \item Kickers Count
  \item Betting gives you two ways to win
\end{itemize}

\section{Old Poker Sayings}

In the comic strip ``The Phantom'', the legend of The Phantom is
spread by forest people through a series of Old Jungle Sayings.
``When the Phantom strikes, lightning stands still'' is one example.
In the same way, here's a list of Old Poker Sayings, which contain
a lot of wisdom which you will do well to heed.

\begin{itemize}
  \item Poker is a battle between a made hand and a drawing hand
  \item Poker is an Information War
  \item Beware of the Speech
  \item To win a tournament you have to win with Ace-King and win
    against Ace-King
  \item You can't bluff a bluffer
  \item Nobody bluffs a made hand
  \item Every hand's a winner and every hand's a loser
  \item Big Blind Special
  \item Never trust a poker player
  \item Send it!
  \item It's easier to bet all-in than it is to call all-in
  \item Poker is war disguised as cards
  \item Nice guys finish last in poker
  \item If it's good enough to call with, it's good enough to raise with.
  \item Don't go broke with one pair
\end{itemize}

You might know more Old Poker Sayings that belong in this list;
email me any you think I'm missing and I'll add them to the next
edition. There's always room for more Old Poker Sayings.

% FIXME: Shorten this; too much GUFF
One famous US Presidential saying is actually a poker saying!
US President Harry S. Truman was fond of ``The Buck Stops Here'' and
had it made into a sign on the White House desk. The Buck in
this saying isn't the almighty dollar, it's the dealer button. When
you move the dealer button, you're ``passing the buck'' which became a
metaphor for shifting responsibility to the next guy. As the US
President, Truman had no-one higher in the chain of command to pass
problems up to. The Buck stops here.

If you read the list closely you'd notice that I snuck in
Send it! as an Old Poker Saying. It is. I've been using it for at
least fifteen years. If you're going to make it in the big Sydney
games, you'll be hearing it a lot. Send it!
